A subjective Bayesian framework for synthesizing deep uncertainties in climate risk management

### 2022-10-05 papers/earths-future/submission_02.tex
changes made; need to update letter to reviewers

% cite things in supplemental info file
% see https://www.overleaf.com/learn/how-to/Cross_referencing_with_the_xr_package_in_Overleaf
% it's a big strange for Overleaf!
\usepackage{xr-hyper}
\makeatletter
\newcommand*{\addFileDependency}[1]{% argument=file name and extension
  \typeout{(#1)}
  \@addtofilelist{#1}
  \IfFileExists{#1}{}{\typeout{No file #1.}}
}
\makeatother
\newcommand*{\myexternaldocument}[1]{%
    \externaldocument{#1}%
    \addFileDependency{#1.tex}%
    \addFileDependency{#1.aux}%
}
\myexternaldocument{supplemental}

@@

Aging infrastructure and changes in regulations, finance, patterns of population and infrastructure use, and climate challenge the ability of critical infrastructures to meet design objectives \cite{doss-gollin_txtreme:2021,doss-gollin_fatalism:2020,chester_reliable:2020,asce_infrastructure_climate:2021,ho_dams:2017}.
To achieve acceptable performance with reasonable planning efforts, current practice in engineering, infrastructure design, and regulation relies heavily on standards that specify design events or conditions that buildings and infrastructure should safely withstand \cite{bruneau_multihazard:2017}.
For example, \gls{fema}, local governments, and engineering consultants produce local floodplain maps in many communities.
Buildings in the designated floodplain are subject to specific regulations, such as flood insurance requirements as an eligibility requirement for federally backed mortgages \cite{kousky_voucher:2014} or minimum elevations for new construction \cite{asce_24-05:2006,FEMA_p-55:2011}.
For example, many communities require that new construction be elevated at least \SI{1}{ft} above the nominal 100-year flood level \cite{zarekarizi_suboptimal:2020}.
Although this paper focuses on flooding, similar approaches inform mitigation strategies for a wide range of other hazards \cite{asce_7-10:2013}.

Standards-based risk management frameworks have many advantages, including scalability, explainability, and simplicity.
However, the choice of standard is a complex design and policy choice.
Risk-based design and cost benefit analysis \cite{eijgenraam_flooding:2014,vandantzig_dike:1956,xian_elevation:2017} can provide a quantitative basis for comparing possible standards by emphasizing ``a proportionate response to risk, so that the amount invested in risk reduction is in proportion to the magnitude of the risk and the cost-effectiveness with which that risk may be reduced'' \cite{merz_fluvial:2010}.
This provides a formal basis for choices such as protecting hospitals and critical infrastructure to a higher degree than ordinary buildings \cite{asce_7-10:2013}.
However, these methods are silent on how standards should balance trade-offs, not only between cost and performance but also between other stakeholder values such as sense of place, distributive justice, and safety \cite{keller_management:2021,helgeson_manifesto:2022,quinn_rivalframings:2017,bessette_vimm:2017,vezer_epistemic:2018}.

Moreover, estimates of the performance trade-offs require implicit or explicit assumptions about the likelihood of different possible futures.
Current practice emphasizes nominally objective methods that can be applied consistently across locations.
For example, the \gls{usgs} Bulletin 17C specifies procedures for estimating flood frequency \cite{bulletin17c:2019}.
Similarly, \gls{noaa} Atlas 14 provides estimates of the intensity, duration, and frequency of extreme rainfall \cite{atlas14_texas:2018,noaa_nonstationary:2022}.
One statistical assumptions these analyses make is stationarity (the assumption that past and future hazard come from the same \gls{pdf}), but global climate change and local environmental changes have cast scrutiny on this assumption \cite{merz_review:2014,milly_stationarity:2008,doss-gollin_robustadaptation:2019}.
While some methods have been proposed for incorporating nonstationarity into risk analyses \cite<see>[for a review]{Salas:2018ge}, these assume specific forms of a trend which may not adequately represent physical processes or sample only a subset of uncertainties \cite{doss-gollin_robustadaptation:2019,montanari_immortal:2014,serinaldi_undead:2015}.
At least in part because of the challenges associated with developing objective methods to select from diverging projections of future hazard, official guidance continues to rely on the stationarity assumption \cite{bulletin17c:2019,atlas14_texas:2018}.

@@

We proceed as follows.
In \cref{sec:analysis} we present three formal decision analytic frameworks for analyzing an ensemble of \gls{slr} simulations, building through existing approaches for exploratory modeling (\cref{sec:analysis-explore}) and scenario analysis (\cref{sec:analysis-condition}) towards a formal Bayesian method for transparently synthesizing deep uncertainty through subjective prior beliefs (\cref{sec:analysis-synthesize}).
In \cref{sec:case-study} we describe the case study in detail.
Next, in \cref{sec:results}, we present results for each of the three decision lenses and discuss the advantages and limitations of each theoretical approach.
In \cref{sec:caveats} we discuss limitations of the study and future research needs.
Finally in \cref{sec:conclusions} we discuss key findings and implications for policy and practice.

\section{Decision analytic framework}\label{sec:analysis}

In this section we outline our framework for decision analysis under deep uncertainty, maintaining a high level of generality.
In the next we discuss application to the house elevation case study.

Following \cref{fig:flowchart}, we consider using $J$ \glspl{sow}, $\vb{s} = \qty{s_1, s_2, \ldots, s_J}$, to evaluate $I$ candidate decisions, $\vb{x} = \qty{x_1, x_2, \ldots, x_I}$.
For each \gls{sow} $s_j \in \vb{s}$ and decision $x_i \in \vb{x}$ we use a system model $f$ (comprised of multiple components) to calculate a set of metrics describing the performance of decision $x_i$ on \gls{sow} $s_j$, which we denote $u_{ij} = f(x_i, s_j)$.
While we assume for simplicity that the decision space is known and finite, this approach could be coupled to a policy search model that proposes candidate decisions.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{bayes-rdm.pdf}
  \caption{
    Outline of the proposed decision-analytic framework.
    In \cref{sec:analysis-explore} we use an exploratory framework to quantify the performance (c) of candidate decisions (a) under a large ensemble of possible futures (b).
    In \cref{sec:analysis-condition} we illustrate the ``multiple PDF problem'' by creating probability distributions over outcomes (e) that are conditional upon specific models describing the likelihood of different futures (d).
    In our case study, these are \gls{rcp} scenarios and physical models of sea level rise.
    Finally in \cref{sec:analysis-synthesize} we propose a subjective Bayesian framework for synthesizing across deep uncertainties by re-weighting sampled futures (f).
  }\label{fig:flowchart}
\end{figure}

\subsection{Explore}\label{sec:analysis-explore}

A first analytical step is to use the model in an ``exploratory'' mode.
Exploratory modeling strives to avoid making explicit assumptions about the likelihood of different \glspl{sow} and instead seeks to generate new knowledge \cite{bankes:1993} by systematically exploring a large number of possible futures, emphasizing interactions between different uncertainties \cite{reed_msdbook:2022}.
Exploratory modeling is often paired with analyses that identify relevant scenarios  \cite{lamontagne_discovery:2018,groves_scenarios:2007} or summarize a system's response to forcing \cite{Poff:2015jn,Steinschneider:2015kk,sriver_sealevel:2018}.
Despite the aversion to strong assumptions about the likelihood of different futures, subjective modeling decisions such as the choice of system model, the set of candidate decisions, the criteria used to assess outcomes, and the choice of how to sample \glspl{sow} can strongly influence results \cite{quinn_exploratory:2020}.

\subsection{Condition}\label{sec:analysis-condition}

Although exploratory modeling is a useful framework for understanding systems, there are many questions that it cannot answer.
For example, answering questions like ``what is the 95th percentile of metric $u$ under decision $x$''  or ``what is the probability of exceeding a critical threshold'' requires estimating the probability distribution over outcomes \cite<see>[for a general discussion]{schneider_scenarios:2002}.

One way to interpret an ensemble of \glspl{sow} is as \gls{iid} draws from some true data generating process.
This commonly arises when a single deep uncertainty (\eg, an emissions pathway) is used as an input for a fully probabilistic model; we call this scenario-conditional probabilistic analysis.
We add this concept to our framework with boxes (d) and (e) of \cref{fig:flowchart}, denoting the particular scenario (\ie, the assumed input) $M_k$.
If the \glspl{sow} are drawn \gls{iid} from $M_k$, the set of outcomes $u_{i, j}$ can be interpreted as \gls{iid} draws from the conditional distribution over outcomes, $p(u | x_i, M_k)$, allowing estimation of descriptive metrics.
This approach allows for quantification of uncertainty \emph{within} models of deeply uncertain processes, but can only qualitatively characterize uncertainty \emph{between} models \cite{wong_nola:2017,ruckert_coastal:2019,sharma_rcp:2021}.

We can also use this approach to understand \gls{dmdu} methods that sample a set of parameters from fixed ranges.
For example, \citeA{sriver_sealevel:2018} sample parameters describing the rate of \gls{slr} across a range of values to inform coastal adaptation.
Similarly, \citeA{trindade_waterpathways:2020} checks the performance of candidate decisions against an ensemble of synthetic time series of streamflow, water demand, and other parameters by sampling parameters that transform the available data over a plausible range \cite<\ie, robustness metrics; see>[for details]{mcphail_robustness:2019,herman:2015}.
Since sampling over a range is equivalent to sampling from a Uniform distribution, this assumption is equivalent to assuming the true inputs to come from independent Uniform distributions (one for each parameter) bounded by the plausible range.
Many limitations of Uniform and other noninformative priors have been documented in the literature, including (i) that they can induce unrealistic (and non-Uniform) priors over functions of parameters and (ii) results can change when the parameterization of a given process is altered \cite{gupta_prior:2022,seaman_noninformative:2012}.

Our primary concern is not that subjective assumptions about the likelihood of different futures are wrong -- this is, almost surely, inevitable -- but that when these assumptions are opaque and presented without critique or validation they may lead to inscrutable decision processes and sub-optimal decisions.

\subsection{Synthesize}\label{sec:analysis-synthesize}

As discussed in the previous subsection, estimating the expectation of arbitrary functions under uncertainty is a critical component of decision making under uncertainty.
If the \glspl{sow} are drawn \gls{iid} from the target distribution, then the expected value of some function $f(\mathbf{s})$, can be readily estimated following the Monte Carlo estimate $\mathbb{E}\qty[f(\vb{s})] \approx \frac{1}{N} \sum_{j=1}^N f(\vb{s}_j)$.
However, this is often not the case; for example, a fixed set of simulations not drawn from the target distribution may be available for analysis.
In this case, the formula may be adjusted to a weighted Monte Carlo estimate:
\begin{equation}
  \mathbb{E}\qty[f(\vb{s})] \approx \sum_{i=j}^N w_j f(\vb{s}_j),
\end{equation}
where $\sum_{j=1}^J w_j = 1$.

The challenge then becomes to suitably estimate the $w_j$.
Paralleling joint probability methods for statistical analysis of tropical cyclones, we use a grid-based approach \cite{johnson_clara:2013,resio_probabilities:2007,toro_jpm-os:2010}.
First, we project the \glspl{sow} $\vb{s} \in \Omega$ onto a low-dimensional representation, which we denote $\qty{\psi_1, \psi_2, \ldots, \psi_J} \in \Psi$.
Then, we partition the parameter space into a region corresponding to each \gls{sow} and integrating the \gls{pdf} $p_\mathrm{belief}$ over each region.

We present here the case where the $\psi_j$ are one-dimensional; extensions to higher dimensions are possible.
We first sort the $\psi_j$  from least to greatest so that $\psi_{j-1} \leq \psi_j$, ($j \neq 1$).
Defining a cumulative distribution function
$F_\mathrm{belief} \qty(\psi) = \int_{-\infty}^{\psi} p_\mathrm{belief}(\psi') \dd{\psi'}$,
we calculate weights as
\begin{equation}\label{eq:weights}
  w_j = \begin{cases}
    F_\mathrm{belief}\qty(\frac{\psi_1 + \psi_2}{2})                                                              & j = 1     \\
    F_\mathrm{belief}\qty(\frac{\psi_{j} + \psi_{j+1}}{2}) - F_\mathrm{belief}\qty(\frac{\psi_{j-1} + \psi_j}{2}) & 1 < j < J \\
    1 - F_\mathrm{belief}\qty(\frac{\psi_{J-1} + \psi_J}{2})                                                      & j = J.
  \end{cases}
\end{equation}
This step is illustrated in \cref{fig:grid-sketch}.
Diagnostic checks, such as examining the histogram of weights (not shown), may be valuable protections against degeneracy.

@@

The aim of this re-weighting framework is to integrate an ensemble of \glspl{sow} used for exploratory modeling into formal decision analysis, even when the \glspl{sow} deliberately over- or under-sample some regions of the parameter space.
As in \cref{sec:analysis-condition}, we must condition on a model: where the analysis of \cref{sec:analysis-condition} conditions upon deep uncertainties, the approach outlined in this subsection synthesizes across them.
We reiterate that stakeholders and experts will not, in general, agree on $p_\mathrm{belief}$ because there is not, even conceptually, an objectively correct choice \cite{oreskes_verification:1994,walker_deep:2013}.
However, we posit that since we cannot be ``right,'' it is valuable to maximize the transparency of our implicit probabilistic assumptions, and suggest that writing down an explicit model for $p_\mathrm{belief}$ supports this aim.

Considering multiple probabilistic models for $p_\mathrm{belief}$ can also be useful for understanding the sensitivity of the decision to the choice of $p_\mathrm{belief}$.
Further, the sensitivity, or lack thereof, of different objectives to the choice of $p_\mathrm{belief}$ may be useful for identifying future research needs.

@@

To illustrate the proposed decision analytic framework, we model a one-time decision of whether to elevate a house, and if so by how much (\cref{fig:xlrm}).
Following the approach outlined in \citeA{zarekarizi_suboptimal:2020}, we focus on a case study of a \emph{hypothetical} house in Norfolk, VA.
For interpretability, we focus on deep uncertainty in \gls{msl} and treat storm surge and other model parameters as shallow uncertainties as shown in \cref{tab:uncertainties}.
We use the notation developed in the previous section to describe the case study.
Specifically,
\begin{enumerate}
  \item The decision vector $\vb{x}$ is comprised of discrete possible house heightenings ($\Delta h$); we consider $\Delta h = \qty{\SI{0}{ft}, \SI{0.25}{ft}, \ldots, \SI{12}{ft}}$.
  \item The \glspl{sow} describe annual time series of \gls{msl} over the $T=70$ year house lifetime so $\vb{s} \in \mathbb{R}^T$
  \item The system model $f$ quantifies up-front costs (the cost of elevating) and lifetime expected damages (the structural cost of experiencing floods), given a decision $x_i$ and \gls{sow} $s_j$, by integrating economic and engineering damage models over a probability distribution for storm surge. We elaborate upon these metrics in \cref{sec:case-metrics}.
\end{enumerate}
In the remainder of this section we describe data sources and treatment of \gls{slr} (\cref{sec:case-slr}), storm surge (\cref{sec:case-surge}), damages and metrics (\cref{sec:case-metrics}), and finally the subjective priors $p_\mathrm{belief}$ used to apply the re-weighting method described in \cref{sec:analysis-synthesize} to this case study (\cref{sec:case-priors}).

@@

We analyze simulations of \gls{msl} at Sewells Point, VA from four probabilistic physical models using data published in \citeA{ruckert_coastal:2019}.
The four models considered are (i) the BRICK model (version 0.2) with slow (``BRICK Slow'') and (ii) fast (``BRICK Fast'') ice sheet dynamics \cite{wong_brick0.2:2017}, (iii) the \citeA{kopp_probabilistic:2014} model (``K14''), and (iv) the \citeA{deconto_antarctica:2016} model (``DP16'').
The \citeA{kopp_probabilistic:2014} and \citeA{deconto_antarctica:2016} models have a ten year time step, which we linearly interpolate onto a one year time step for consistency.
These four models represent physical processes, particularly of ice sheet dynamics, in different ways.
The differences in these models leading to diverging estimates of how \gls{msl} responds to temperature.
For a discussion of these model outputs we refer the reader to \citeA{ruckert_coastal:2019}.

Estimates of nonstationary \gls{msl} also depend on anthropogenic forcing, which is itself deeply uncertain \cite{ho_scenarios:2019,srikrishnan_probabilistic:2022}.
To sample this uncertainty, we use simulations from each physical model under four \gls{rcp} scenarios, yielding sixteen time-varying distributions of \gls{msl}.

@@

We use a Gamma distribution for all three priors, parameterized following \cref{eq:gamma-dist}.
The distributions were chosen to be illustrative, rather than to reflect any particular scientific consensus.
The Gamma distribution is a flexible distribution that can be used to model skewed, lower-bounded distributions, making it an appropriate choice for modeling subjective uncertainty about \gls{slr}.
\Cref{tab:slr-priors} specifies the parameters of these distributions, as well as some quantiles of the distributions.
Their \glspl{pdf} are also plotted in \cref{fig:lsl-priors-weights}(A).

>>>>

@@

Aging infrastructure and changes in regulations, finance, patterns of population and infrastructure use, and climate challenge the ability of critical infrastructures to meet design objectives \cite{doss-gollin_txtreme:2021,doss-gollin_fatalism:2020,chester_reliable:2020,asce_infrastructure_climate:2021,ho_dams:2017}.
To achieve acceptable performance with reasonable planning efforts, current practice in engineering, infrastructure design, and regulation relies heavily on standards that specify design events or conditions that buildings and infrastructure should safely withstand \cite{bruneau_multihazard:2017}.
For example, \gls{fema}, local governments, and engineering consultants produce local floodplain maps in many communities.
Buildings in the designated floodplain are subject to specific regulations, such as flood insurance requirements as an eligibility requirement for federally backed mortgages \cite{kousky_voucher:2014} or minimum elevations for new construction \cite{asce_24-05:2006,FEMA_p-55:2011}.
Although this paper focuses on flooding, similar approaches inform mitigation strategies for a wide range of other hazards \cite{asce_7-10:2013}.

Standards-based risk management frameworks have many advantages, including scalability, explainability, and simplicity.
However, the choice of standard is a complex design and policy choice.
Risk-based design and cost benefit analysis \cite{eijgenraam_flooding:2014,vandantzig_dike:1956,xian_elevation:2017} offer a quantitative framework for comparing possible standards by emphasizing ``a proportionate response to risk, so that the amount invested in risk reduction is in proportion to the magnitude of the risk and the cost-effectiveness with which that risk may be reduced'' \cite{merz_fluvial:2010}.
This provides a formal basis for choices such as protecting hospitals and critical infrastructure to a higher degree than ordinary buildings \cite{asce_7-10:2013}.
However, these methods are silent on how standards should balance trade-offs, not only between cost and performance but also between other stakeholder values such as sense of place, distributive justice, and safety \cite{keller_management:2021,helgeson_manifesto:2022,quinn_rivalframings:2017,bessette_vimm:2017,vezer_epistemic:2018}.

Moreover, estimates of performance trade-offs require implicit or explicit assumptions about the likelihood of different possible futures.
Current practice emphasizes nominally objective methods that can be applied consistently across locations.
For example, the \gls{usgs} Bulletin 17C specifies procedures for estimating flood frequency \cite{bulletin17c:2019}.
Similarly, \gls{noaa} Atlas 14 provides estimates of the intensity, duration, and frequency of extreme rainfall \cite{atlas14_texas:2018,noaa_nonstationary:2022}.
One statistical assumption these analyses make is stationarity (the assumption that past and future hazard come from the same \gls{pdf}), but global climate change and local environmental changes have cast scrutiny on this assumption \cite{merz_review:2014,milly_stationarity:2008,doss-gollin_robustadaptation:2019}.
While some methods have been proposed for incorporating nonstationarity into risk analyses \cite<see>[for a review]{Salas:2018ge}, these assume specific forms of a trend which may not adequately represent physical processes or sample only a subset of uncertainties \cite{doss-gollin_robustadaptation:2019,montanari_immortal:2014,serinaldi_undead:2015}.
At least in part because of the challenges associated with developing objective methods to select from diverging projections of future hazard, official guidance continues to rely on the stationarity assumption \cite{bulletin17c:2019,atlas14_texas:2018}.

@@

We proceed as follows.
In \cref{sec:analysis} we present three formal decision analytic frameworks for analyzing an ensemble of \gls{slr} simulations, building through existing approaches for exploratory modeling scenario analysis to identify a need for synthesizing across scenarios.
We present a formal framework for transparently synthesizing deep uncertainties in \cref{sec:reweighting}.
In \cref{sec:case-study} we describe the didactic case study.
Next, in \cref{sec:results}, we present results for each of the three decision lenses and discuss the advantages and limitations of each theoretical approach.
In \cref{sec:caveats} we discuss limitations of the study and future research needs.
Finally in \cref{sec:conclusions} we discuss key findings and implications for policy and practice.

\section{Conceptual framework}\label{sec:analysis}

In this section we introduce a conceptual framework and notation for decision analysis under deep uncertainty.
We then apply this framework to understand exploratory modeling (\cref{sec:analysis-explore}) and scenario-conditional probabilistic analysis (\cref{sec:analysis-condition}).
In \cref{sec:analysis-synthesize} we discuss why synthesizing across multiple scenarios is necessary.
In \cref{sec:reweighting} we provide a formal method for synthesizing across scenarios.

It is common in climate risk management to work with an ensemble of trajectories of climate, economic, and other key variables, which we refer to as \glspl{sow}.
Often these \glspl{sow} are computationally expensive to generate, and thus it is important to use them efficiently.
Following \cref{fig:flowchart}, we consider using $J$ \glspl{sow}, $\vb{s} = \qty{s_1, s_2, \ldots, s_J}$, to evaluate $I$ candidate decisions, $\vb{x} = \qty{x_1, x_2, \ldots, x_I}$.
For each \gls{sow} $s_j \in \vb{s}$ and decision $x_i \in \vb{x}$ we use a system model $f$ (comprised of multiple components) to calculate a set of metrics describing the performance of decision $x_i$ on \gls{sow} $s_j$, which we denote $u_{ij} = f(x_i, s_j)$.
While we assume for simplicity that the decision space is known and finite, this approach could be coupled to a policy search model that proposes candidate decisions.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{bayes-rdm.pdf}
  \caption{
    Outline of the proposed decision-analytic framework.
    In \cref{sec:analysis-explore} we use an exploratory framework to quantify the performance of candidate decisions under a large ensemble of possible futures.
    In \cref{sec:analysis-condition} we illustrate the ``multiple PDF problem'' by creating probability distributions over outcomes that are conditional upon specific probabilistic scenarios.
    In our case study, these scenarios correspond to combinations of emissions pathways with physical models for sea level rise.
    Then in \cref{sec:analysis-synthesize} we describe the need for synthesizing insight across scenarios.
    Finally in \cref{sec:reweighting} we provide a formal framework for doing so.
  }\label{fig:flowchart}
\end{figure}

\subsection{Exploratory modeling}\label{sec:analysis-explore}

A first analytical step is to use the model in an ``exploratory'' mode.
Exploratory modeling strives to avoid making explicit assumptions about the likelihood of different \glspl{sow} and instead seeks to generate new knowledge \cite{bankes:1993} by systematically exploring a large number of possible futures, emphasizing interactions between different uncertainties \cite{reed_msdbook:2022}.
Exploratory modeling is often paired with analyses that identify relevant scenarios  \cite{lamontagne_discovery:2018,groves_scenarios:2007} or summarize a system's response to forcing \cite{Poff:2015jn,Steinschneider:2015kk,sriver_sealevel:2018}.
Despite the aversion to strong assumptions about the likelihood of different futures, subjective modeling decisions such as the choice of system model, the set of candidate decisions, the criteria used to assess outcomes, and the choice of how to sample \glspl{sow} can strongly influence results \cite{quinn_exploratory:2020,quinn_rivalframings:2017,moallemi_decisionsupport:2020}.

\subsection{Scenario-conditional probabilistic analysis}\label{sec:analysis-condition}

Although exploratory modeling is a useful framework for understanding systems, there are many questions that it cannot answer.
For example, answering questions like ``what is the 95th percentile of metric $u$ under decision $x$''  or ``what is the probability of exceeding a critical threshold'' requires an implicit or explicit probability distribution over outcomes \cite<see>[for a general discussion]{schneider_scenarios:2002}.

One way to interpret an ensemble of \glspl{sow} is as {iid} draws from some probabilistic data generating process.
This commonly arises when a single deep uncertainty (\eg, an emissions pathway) is used as an input for a stochastic model.
We illustrate this concept in boxes (d) and (e) of \cref{fig:flowchart}, denoting the particular scenario $M_k$.
By treating the \glspl{sow} as \gls{iid} draws from $M_k$, the set of outcomes $u_{i, j}$ can be interpreted as \gls{iid} draws from the conditional distribution over outcomes, $p(u | x_i, M_k)$.
This ``scenario-conditional'' probabilistic interpretation of \glspl{sow} allows for fully probabilistic quantification of uncertainty and optimization, conditional on a particular scenario.
For example, \citeA{fletcher_learning:2019} uses stochastic dynamic programming to quantify the value of flexibility in water resources planning.
However, only a single \gls{rcp} scenario (\gls{rcp} 8.5) is used.
While the analysis could be repeated for other \gls{rcp} scenarios, the scenario-conditional analysis framework only qualitatively characterize uncertainty \emph{between} scenarios \cite{wong_nola:2017,ruckert_coastal:2019,sharma_rcp:2021}.

We can also characterize \gls{dmdu} methods that sample a set of parameters from fixed ranges as scenario-conditional analyses.
For example, \citeA{sriver_sealevel:2018} sample parameters describing the rate of \gls{slr} across a range of values to inform coastal adaptation.
Similarly, \citeA{trindade_waterpathways:2020} checks the performance of candidate decisions against an ensemble of synthetic time series of streamflow, water demand, and other parameters by sampling parameters that transform the available data over a plausible range.
Analyses that use this methodology are implicitly assuming a single probabilistic model in which different variables are drawn from independent Uniform distributions.
Many limitations of Uniform and other noninformative priors have been documented in the literature, including (i) that they can induce unrealistic implicit priors over functions of parameters and (ii) results are sensitive to the parameterization of a given process \cite{seaman_noninformative:2012}.
Yet while replacing Uniform distributions with alternatives such as maximum-entropy distributions can address some of these challenges \cite<\eg,>{gupta_prior:2022}, subjective modeling choices remain necessary.
Our primary concern here is not that these subjective modeling assumptions are wrong -- this is, almost surely, inevitable -- but that when these assumptions are opaque and presented without critique or validation \cite<see>[regarding the importance of iterative critique]{gelman_workflow:2020} they may lead to inscrutable decision processes and poor decisions.

\subsection{Synthesizing deep uncertainties for decision analysis}\label{sec:analysis-synthesize}

Scenario-conditional probabilistic analysis allows for uncertainty quantification and optimization, and is valuable in many copntexts.
However, scenarios are often explicitly provided without probabilities or likelihoods \cite<\eg, the \acrlongpl{ssp}>{vanvuuren_conditional:2008}.
Thus, any such analysis is silent on the question of how to combine information across different scenarios.
We term this the ``multiple \gls{pdf} problem''.
Decision making around the multiple \gls{pdf} problem is susceptible to the cognitive biases that interfere with decision-making under uncertainty more generally \cite{tversky_judgment:1974,morgan_uncertainty:1990,srikrishnan_probabilistic:2022}.
For example, while many analyses treat all scenarios as equally likely, this is often inconsistent with available information and can lead to poor decisions and outcomes \cite{wigley_equallylikely:2001,ho_scenarios:2019,hausfather_scenarios:2020}.
Other analyses suggest using the worst-case scenario as a conservative measure.
However, this approach is also problematic, since (i) there are no fundamental limits on what constitutes a worst-case scenario and (ii) improving performance under unlikely worst-case scenarios may lead to substantially impaired performance under more likely scenarios, which may or may not be acceptable to relevant stakeholders.
There is thus a critical need for synthesizing insights across multiple probabilistic scenarios.

\section{Re-weighting SOWs to synthesize across scenarios}\label{sec:reweighting}

In this section we provide a formal method for synthesizing across multiple scenarios to inform climate risk management.
Our objective is to develop a framework that (i) is conceptually and practically amenable to exploratory modeling; (ii) makes subjective modeling choices explicit and transparent; and (iii) allows decision analysts to estimate a probability distribution over outcomes.

A particular need is to estimate the expectation of functions over \glspl{sow} (\ie, box e in \cref{fig:flowchart}).
If the $J$ \glspl{sow} are drawn \gls{iid} from some distribution that credibly represents the true likelihood of different futures then the expected value of such a function, $f(\mathbf{s})$, can be readily approximated using the Monte Carlo estimate $\mathbb{E}\qty[f(\vb{s})] \approx \frac{1}{N} \sum_{j=1}^N f(\vb{s}_j)$.
However, this is often not the case.
For example, in \cref{sec:case-study} we will consider decision analysis where the \glspl{sow} are sampled from multiple physical models and \gls{rcp} scenarios, considering that not all \gls{rcp} scenarios are equally likely and that not all physical models are equally skillful.
In this case, the formula may be adjusted to a weighted Monte Carlo estimate:
\begin{equation}
  \mathbb{E}\qty[f(\vb{s})] \approx \sum_{i=j}^N w_j f(\vb{s}_j),
\end{equation}
where $\sum_{j=1}^J w_j = 1$.

The challenge then becomes to suitably estimate the $w_j$.
Many such methods exist; drawing from joint probability methods for statistical analysis of tropical cyclones, we employ a grid-based approach \cite{johnson_clara:2013,resio_probabilities:2007,toro_jpm-os:2010}.
First, we project the \glspl{sow} $\vb{s} \in \Omega$ onto a low-dimensional representation, which we denote $\qty{\psi_1, \psi_2, \ldots, \psi_J} \in \Psi$.
Then, we partition the parameter space into a region corresponding to each \gls{sow} and integrating the probability $p(\psi)$ over each region.

Doing so requires a probabilistic model for this low-dimensional representation of the \glspl{sow},  $p(\psi)$.
We denote this model $p_\mathrm{belief}$ to emphasize that it represents a subjective belief about the \glspl{sow}, rather than an objectively verifiable choice.
In general we do not expect that stakeholders and experts will agree on $p_\mathrm{belief}$ because there is not, even conceptually, an objectively correct choice \cite{oreskes_verification:1994,walker_deep:2013}.
However, we posit that since we cannot be ``right,'' it is valuable to maximize the transparency of our implicit probabilistic assumptions, and suggest that writing down an explicit model for $p_\mathrm{belief}$ supports this aim.
Choices for $p_\mathrm{belief}$ can be drawn from many sources, including expert elicitation or  results of previous analyses.
These models can be interpreted as prior beliefs about \gls{slr} that could be incorporated into a Bayesian analysis as additional data is collected in the future, and thus can draw from literature on Bayesian prior selection and prior predictive checks \cite{gelman_workflow:2020}.

We present here the case where the $\psi_j$ are one-dimensional; extensions to higher dimensions are possible.
We first sort the $\psi_j$  from least to greatest so that $\psi_{j-1} \leq \psi_j$, ($j \neq 1$).
Defining a cumulative distribution function
$F_\mathrm{belief} \qty(\psi) = \int_{-\infty}^{\psi} p_\mathrm{belief}(\psi') \dd{\psi'}$,
we calculate weights as
\begin{equation}\label{eq:weights}
  w_j = \begin{cases}
    F_\mathrm{belief}\qty(\frac{\psi_1 + \psi_2}{2})                                                              & j = 1     \\
    F_\mathrm{belief}\qty(\frac{\psi_{j} + \psi_{j+1}}{2}) - F_\mathrm{belief}\qty(\frac{\psi_{j-1} + \psi_j}{2}) & 1 < j < J \\
    1 - F_\mathrm{belief}\qty(\frac{\psi_{J-1} + \psi_J}{2})                                                      & j = J.
  \end{cases}
\end{equation}
This step is illustrated in \cref{fig:grid-sketch}.
By the definition of cumulative distribution functions, $F_\mathrm{belief}(b) - F_\mathrm{belief}(a) = \int_a^b p_\mathrm{belief}(\psi') \dd{\psi'}$.
Diagnostic checks, such as examining the histogram of weights to (not shown), may be useful protections against degeneracy.

@@

The aim of this re-weighting framework is to integrate an ensemble of \glspl{sow} used for exploratory modeling into formal decision analysis, even when the \glspl{sow} deliberately over- or under-sample some regions of the parameter space.
As in \cref{sec:analysis-condition}, we must condition on a model: where the analysis of \cref{sec:analysis-condition} conditions upon deep uncertainties, the approach outlined in this subsection synthesizes across them.
Considering multiple probabilistic models for $p_\mathrm{belief}$ can also be useful for understanding the sensitivity of the decision to the choice of $p_\mathrm{belief}$.
Further, the sensitivity, or lack thereof, of different objectives to the choice of $p_\mathrm{belief}$ may be useful for identifying future research needs.

@@

To illustrate the proposed decision analytic framework, we model a one-time decision of whether to elevate a house, and if so by how much (\cref{fig:xlrm}).
Following the approach outlined in \citeA{zarekarizi_suboptimal:2020}, we focus on a case study of a \emph{hypothetical} house in Norfolk, VA.
For interpretability, we focus on deep uncertainty in \gls{msl} and treat storm surge and other model parameters as shallow uncertainties as shown in \cref{tab:uncertainties}.
We use the notation developed in the previous section to describe the case study.
Specifically,
\begin{enumerate}
  \item The decision vector $\vb{x}$ is comprised of discrete possible house heightenings ($\Delta h$); we consider $\Delta h = \qty{\SI{0}{ft}, \SI{0.25}{ft}, \ldots, \SI{12}{ft}}$.
  \item The \glspl{sow} describe annual time series of \gls{msl} over the $T=70$ year house lifetime so $\vb{s} \in \mathbb{R}^T$
  \item The system model $f$ quantifies up-front costs (the cost of elevating) and lifetime expected damages (the structural cost of experiencing floods), given a decision $x_i$ and \gls{sow} $s_j$, by integrating economic and engineering damage models over a probability distribution for storm surge. We elaborate upon these metrics in \cref{sec:case-metrics}.
\end{enumerate}
In the remainder of this section we describe data sources and treatment of \gls{slr} (\cref{sec:case-slr}), storm surge (\cref{sec:case-surge}), damages and metrics (\cref{sec:case-metrics}), and finally the subjective probabilistic models $p_\mathrm{belief}$ used to apply the re-weighting method described in \cref{sec:analysis-synthesize} to this case study (\cref{sec:case-priors}).

@@

We analyze simulations of \gls{msl} at Sewells Point, VA from four probabilistic physical models using data published in \citeA{ruckert_coastal:2019}.
The four models considered are (i) the BRICK model (version 0.2) with slow (``BRICK Slow'') and (ii) fast (``BRICK Fast'') ice sheet dynamics \cite{wong_brick0.2:2017}, (iii) the \citeA{kopp_probabilistic:2014} model (``K14''), and (iv) the \citeA{deconto_antarctica:2016} model (``DP16'').
The \citeA{kopp_probabilistic:2014} and \citeA{deconto_antarctica:2016} models have a ten year time step, which we linearly interpolate onto a one year time step for consistency.
Estimates of nonstationary \gls{msl} also depend on anthropogenic forcing, which is itself deeply uncertain \cite{ho_scenarios:2019,srikrishnan_probabilistic:2022}.
To sample this uncertainty, we use simulations from each physical model under four \gls{rcp} scenarios, yielding sixteen time-varying probabilistic scenarios of \gls{msl}.

@@

We use a Gamma distribution for all three priors, parameterized following eq.~(S4).
The distributions were chosen to be illustrative, rather than to reflect any particular scientific consensus.
The Gamma distribution is a flexible distribution that can be used to model skewed, lower-bounded distributions, making it an appropriate choice for modeling subjective uncertainty about \gls{slr}.
\Cref{tab:slr-priors} specifies the parameters of these distributions, as well as some quantiles of the distributions.
Their \glspl{pdf} are also plotted in \cref{fig:lsl-priors-weights}(A).
